Analyzing X-ray images in Python with scikit-image

### 2016-08-21 skimage_tomo.tex

Given the large size of tomography datasets, the speed of execution of image
processing operations is a critical issue. \texttt{scikit-image} relies mostly
on calls to \texttt{NumPy} operations, with most of them that are performed in
optimized compiled code (C or Fortran). Performance-critical parts of the code
that cannot call efficient \texttt{NumPy} code are implemented in
\texttt{Cython}. \texttt{Cython} \citep{Behnel2011} is an extension of the
Python language that supports explicit type declarations, and is compiled
directly to C.

However, basic \texttt{scikit-image} code runs on a single core. Computing
workstations and servers used for X-ray imaging typically have several tens of
cores. Parallelization of the computing workflow can be achieved in different
ways. The most trivial parallelization scheme consists in applying the same
workflow to different images, on different cores. However, finer-grained
parallelization is preferable when prototyping the processing workflow.

@@

On a different note, prototyping the image processing workflow requires several trial and error
executions of scripts containing different operations. The \texttt{joblib}
module provides a caching mechanism that avoids the repetition of function
calls, if their arguments have not changed:
\begin{lstlisting}
>>> from skimage import filters, data
>>> from joblib import Memory
>>> mem = Memory(cachedir='/tmp/joblib')
>>> median = mem.cache(filters.median)
>>> im = data.camera()
>>> filtered_im = median(im, np.ones((3, 3)))
__________________________
[Memory] Calling skimage.filters.rank.generic.median...
median(array([[156, ..., 152],
       ..., 
       [121, ..., 111]], dtype=uint8), array([[ 1.,  1.,  1.],
       [ 1.,  1.,  1.],
       [ 1.,  1.,  1.]]))
__________________________median - 0.0s, 0.0min
>>> filtered_again = median(im, np.ones((3, 3)))
>>> # the above call did not trigger an evaluation
\end{lstlisting}

Finally, let us note that the biggest performance improvements often come from
the improvement of algorithms (as opposed to computing architectures only). For
example, non-local means denoising \citep{Buades2005} is a costly operation,
since it requires several nested loops, on all pixels and on neighboring
patches to be compared with the pixel-centered patch. Thanks to the
implementation of a more recent algorithm \citep{Darbon2008} that modifies the
internal organization of loops, it was possible to gain a factor of 5 to 10.
The large size of the \texttt{scikit-image} community ensures that possible
algorithmic improvements are discussed regularly.

>>>>

Given the large size of tomography datasets, the execution speed of image
processing operations is of critical concern. \texttt{scikit-image} relies mostly
on calls to \texttt{NumPy} operations, of which most are performed in
optimized compiled code (C or Fortran). Performance-critical parts of \texttt{scikit-image}
that cannot call efficient \texttt{NumPy} code are implemented in
\texttt{Cython}. \texttt{Cython} \citep{Behnel2011} is an extension of the
Python language that supports explicit type declarations, and is compiled
directly to C.

However, basic \texttt{scikit-image} code runs on a single core. Computing
workstations and servers used for X-ray imaging typically have several tens of
cores. Parallelization of the computing workflow can be achieved in multiple
ways. The most trivial parallelization scheme consists of applying the same
workflow to different images, on different cores. However, finer-grained
parallelization is preferable when prototyping the processing workflow.

@@

Another situation that often arises is that, while prototyping a
workflow, scripts (containing the image processing pipeline) are run
several times to experiment with parameters. The \texttt{joblib}
module provides a caching mechanism that avoids the repetition of function
calls if their arguments have not changed:
\begin{lstlisting}
>>> from skimage import filters, data
>>> from joblib import Memory
>>> mem = Memory(cachedir='/tmp/joblib')
>>> median = mem.cache(filters.median)
>>> im = data.camera()
>>> filtered_im = median(im, np.ones((3, 3)))
__________________________
[Memory] Calling skimage.filters.rank.generic.median...
median(array([[156, ..., 152],
       ..., 
       [121, ..., 111]], dtype=uint8), array([[ 1.,  1.,  1.],
       [ 1.,  1.,  1.],
       [ 1.,  1.,  1.]]))
__________________________median - 0.0s, 0.0min
>>> filtered_again = median(im, np.ones((3, 3)))
>>> # the above call did not trigger another evaluation
\end{lstlisting}

Finally, let us note that the biggest performance improvements often come from
the improvement of algorithms (as opposed to computing architectures only). For
example, non-local means denoising \citep{Buades2005} is a costly operation,
since it requires several nested loops, on all pixels and on neighboring
patches to be compared with the pixel-centered patch. Thanks to the
implementation of a more recent algorithm \citep{Darbon2008} that modifies the
internal organization of loops, it was possible to improve execution
time by a factor of roughly ten times.
The large size of the \texttt{scikit-image} community makes it likely for
algorithmic improvements to be discussed regularly.
